Replication on multicore considering hierarchy and characteristics of the network

### 2013-03-14 paper.tex
Some first measurements

\documentclass{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage[colorlinks=true]{hyperref}

@@

% tikz stuff
\usepackage{tikz}
\usetikzlibrary{shapes,positioning,calc}

@@

\paragraph{What do you want to enable?} Increase algorithm performance
on multicores by applying distributed principles, particularly
replication. The challenge then is to guarantee consistency. Previous
work was using primary copy protocols to do that~\cite{Wiesmann2000}, a
passive replication scheme (i.e.\ nodes do not have to contact all
replicas). They do that because they claim that it is not suitable for
clients to know about replication details.

\paragraph{What problem are you solving, and why is it hard?} Three problems:
\begin{description}
\item[placement] machine dependent
\item[consistency] need to consider machine \emph{characteristics};
  multicores are \emph{different} from classical distributed systems.
\item[machine characteristics] diverse, fast changing, complex
\end{description}

@@

>>>>

\documentclass{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage{amssymb}
\usepackage[colorlinks=true]{hyperref}

@@

% tikz stuff
\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{shapes,positioning,calc}

@@

\paragraph{What do you want to enable?} Increase algorithm performance
on multicores by applying distributed principles, particularly
replication. The challenge then is to guarantee consistency. Previous
work was using primary copy protocols to do that~\cite{Wiesmann2000}, a
passive replication scheme (i.e.\ nodes do not have to contact all
replicas). They do that because they claim that it is not suitable for
clients to know about replication details. An atomic broadcast is
often used as foundation of such algorithms. It guarantees in reliable
delivery of messages to every node.

\paragraph{What problem are you solving, and why is it hard?} Three problems:
\begin{description}
\item[placement] machine dependent
\item[consistency] need to consider machine \emph{characteristics};
  multicores are \emph{different} from classical distributed systems.
\item[machine characteristics] diverse, fast changing, complex. One
  example is the batch size for sending a broadcast.
\end{description}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mechanisms}

\subsubsection{Aggregation}

Similarly to what has been done in wireless sensor networks (where it
also matters to reduce the number of messages, but other reasons:
power consumption), we can do aggregation in nodes. In difference to
traditional distributed systems, this works, because it is easy to
deploy custom software on every node in the network. Classical
distributed systems do not typically allow this. Furthermore, reducing
the number of messages at the price of higher complexity does not make
sense there.

Examples for aggregation: number of nodes agreeing to something, find
capabilities (concatenate core ids).

Open questions: How does this work with several senders (everyone is
potentially a sender). Maybe something like in navigation
systems. Every node has an entry point, which forwards the
messages. Nodes are then clustered \ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Implementation}

Build a tree. Every node has parents and children (except for the root
and the leafs). The number of children does not really matter for the
algorithm. In the init phase, we construct the tree (e.g. broadcast:
one root node, all other nodes are children of the root node,
multicast: every node has two children, NUMA tree: two layers, two
different implementations?

Reducing the number of children also helps in select. It reduces the
number of ``channels'' to poll, and therefore the latency of detecting
messages. Instead of

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental results}

on revision 1190:a787513fd28c, experiments for sequential and batch
broadcast working on nos5 and appenzeller

\begin{table}[htb]
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    broadcast algorithm & \multicolumn{2}{c}{nos5} & \multicolumn{2}{c}{gruyere} \\
      & cycles & error & cycles & error \\
    \midrule
    sequential &  4096.1 &  105.2 & 136751.9 &   3061.5 \\
    batch      &  2408.0 &  181.9 &  57360.8 &   5271.5 \\
    tree\footnote{This only works for the first 200 runs, after that,
      performance breaks down, don't know why, but the first 200 runs
      are quite stable}%
               &  \sout{1471.8} &  \sout{187.9} &          &          \\
    \bottomrule
  \end{tabular}
  \caption{Broadcast measurements}
  \label{tab:bc_measurements}
\end{table}

\paragraph{Measurements to do next} %
\begin{itemize}
\item Increase number of cores to talk to in every round. Use a
  separate waitsets, so that we don't have additional overhead from
  the bindings not used in each concrete measurement.
\item Play with batching. Incrementally increase the number of
  messages send in a batch. I assume this is machine specific, which
  means, that this is something programmers do not want to deal
  with. We measure this at dynamically at run-time and find the right
  batch size.
\end{itemize}

\input{gruyere_bc_seq}

\begin{itemize}
\item Move server to a core != 0
\item Use low-level UMP stuff
{
\renewcommand{\labelitemi}{\checkmark}
\item disable yielding set 
}
\item some other item
\item run several request in parallel, give them IDs, can use in the
  measurement struct. Good for everything having transactions (DB,
  transactional memory, consistency in replication systems)
\end{itemize}
